---
title: "Reproducible Research-Project 1"
author: "Shadi"
date: "1/26/2022"
output: github_document
---

## Loading and preprocessing the data
We start by reading the data and looking at the top row of it to find out if any intervention is necessary 
```{r, echo=TRUE}
activity <-read.csv(file="data/activity.csv")
head(activity)
tail(activity)

```

We can observe that both 10 rows at the begining and the end of the data set have NA for steps value to make sure that the this is just not an error, in our next step we tabulate the values for steps 

```{r, echo=TRUE}
head(table(activity$steps))

```
Before moving to the next step we also want to make sure that variable format are consistent for our analysis. 

```{r, echo=TRUE}
str(activity)
```

Apparently date here is coded as factor we have to fix date to be formatted as date.

```{r, echo=TRUE}
activity$date<- as.Date(activity$date)
str(activity$date)
```

Data is now ready for further analysis.

## What is mean total number of steps taken per day?

### Calculation number of step per day

I would like to calculate the total steps in two diffrent way 

* first way 

```{r, echo=TRUE}
totalstep<-tapply(activity$steps,activity$date,sum)
head(totalstep)
```
The result of this code is an array of numbe while the result of the second way code is a data frame stepbydate with two variable of date and total step 
* second way 

```{r, echo=TRUE}
library(dplyr)
Stepbyday<- activity %>% group_by(date) %>% summarise(totalstep=sum(steps))
head(Stepbyday)
```

1.Histogram of the total number of steps taken each day

```{r, echo=TRUE}
hist(Stepbyday$totalstep,main="Histogram of the total number of steps taken each day", xlab="Total step", col = "Yellow")

```

2.Mean and median total number of steps taken per day


``` {r, echo=TRUE}
mean(totalstep,na.rm=TRUE)
median(totalstep,na.rm=TRUE)
``` 
## What is the average daily activity pattern?

1.Make a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)

### Creating the data frame needed for plot time-series

 a. find average number of step per day for y-axis of plot
 
 
```{r, echo=TRUE}
avgstep<- activity %>% group_by(interval) %>% na.omit()  %>% summarise(totalstep=mean(steps))
head(avgstep)
```
 
 b. create a time series graph

```{r, echo=TRUE}
with(avgstep,plot(interval,totalstep,type = "l",  main = "The average daily activity pattern"))
```





2. Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?

``` {r, echo=TRUE}
avgstep[which.max(avgstep$totalstep),]

```


## Imputing missing values

1. Calculate and report the total number of missing values 

```{r, echo=TRUE}
sum(is.na(activity$steps))

```

 2. Devise a strategy for filling in all of the missing values in the datase
 
As a strategy I will replace NA with the average of steps of the day
 
```{r, echo=TRUE}

meansteps=mean(activity$step,na.rm=TRUE)

```
 
3. Create a new dataset that is equal to the original dataset but with the missing data filled in.


To prevent any problem I will create a copy of activity data set and impute NA on that one  
 
 
```{r, echo=TRUE}

stepdata<-activity

stepdata$steps<-replace(stepdata$steps, is.na(stepdata$steps), meansteps)
        

```
 
 checkin if the new dataset have missing values
 
```{r, echo=TRUE}
sum(is.na(stepdata$steps))

```
 

 4. Calculate Histogram,mean,median total number of step per day  for new data.frame
 
```{r, echo=TRUE}

library(ggplot2)
newtotalstep<-tapply(stepdata$steps,stepdata$date,sum)

newtotalstep<- stepdata %>% group_by(date) %>% summarise(totalstep=sum(steps))

hist(newtotalstep$totalstep, xlab = "number of steps",
      main = "total number of steps taken per day")
```

The new mean and median are as follow 


```{r, echo=TRUE}

mean(newtotalstep$totalstep,na.rm=TRUE)
median(newtotalstep$totalstep,na.rm=TRUE)

```

Next, we caclute the diffrence between the new and old mean and median 

```{r, echo=TRUE}

meandif<- mean(totalstep,na.rm=TRUE)-mean(newtotalstep$totalstep,na.rm=TRUE)
meddif<-median(totalstep,na.rm=TRUE)-median(newtotalstep$totalstep,na.rm=TRUE)

print("diffrence in mean is", meandif)

```

#### As  we can see the mean is the same as when we didn't substitue the NA variable,
#### but the median has slightly changed. our new median is equall to are mean.

## Are there differences in activity patterns between weekdays and weekends?


1. Create new  variable in the new data set with filled in missing value.

  {r, echo=TRUE}
week<-weekdays(as.POSIXlt(activity$date))
week<-ifelse ((week =="Sunday" | week =="Saturday") ,"weekend", "weekday") 
newactivity<-data.frame(cbind(newactivity,week))
 
 2. Make a panel plot containing a time series plot 
 a. build the  new variable avrage step per day
 
 {r, echo=TRUE}
  avgsteptaken<-tapply(newactivity$steps,newactivity$interval,mean,na.rm=TRUE)
  newactivity<-data.frame(cbind(avgsteptaken, newactivity))
`

 b. two time series graph  for weekday no weekend 

{r, echo=TRUE}

png("Plots/plot3.png", width=600, height=600) 


par(mfrow=c(2,1))
plot(newactivity$interval, subset(newactivity,week="weekdat")$avgsteptaken,type = "l") 
plot(newactivity$interval, subset(newactivity,week="weekend")$avgsteptaken,type = "l")
