---
title: "Reproducible Research-Project 1"
author: "Shadi"
date: "1/26/2022"
output: github_document
---

## Loading necessary libraries

```{r, echo=TRUE}
library(dplyr)
library(ggplot2)

```

## Loading and preprocessing the data
We start by reading the data and looking at the top row of it to find out if any intervention is necessary 
```{r, echo=TRUE}
activity <-read.csv(file="data/activity.csv")
head(activity)
tail(activity)

```

We can observe that both 10 rows at the begining and the end of the data set have NA for steps value to make sure that the this is just not an error, in our next step we tabulate the values for steps 

```{r, echo=TRUE}
head(table(activity$steps))

```
Before moving to the next step we also want to make sure that variable format are consistent for our analysis. 

```{r, echo=TRUE}
str(activity)
```

Apparently date here is coded as factor we have to fix date to be formatted as date.

```{r, echo=TRUE}
activity$date<- as.Date(activity$date)
str(activity$date)
```

Data is now ready for further analysis.

## What is mean total number of steps taken per day?

### 1. Making a Histogram of the total number of steps taken each day

I would like to start by calculating the total steps in two diffrent way 

* first way 

```{r, echo=TRUE}
totalstep<-tapply(activity$steps,activity$date,sum)
head(totalstep)
```
The result of this code is an array of numbe while the result of the second way code is a data frame stepbydate with two variable of date and total step 

* second way 

```{r, echo=TRUE}
library(dplyr)
Stepbyday<- activity %>% group_by(date) %>% summarise(totalstep=sum(steps))
head(Stepbyday)
```

Then using the caluclated total step I create the histogram plot


```{r, echo=TRUE}
hist(Stepbyday$totalstep,main="Histogram of the total number of steps taken each day", xlab="Total step", col = "Yellow")

```

### 2. Mean and median total number of steps taken per day


``` {r, echo=TRUE}
mean(totalstep,na.rm=TRUE)
median(totalstep,na.rm=TRUE)
``` 
## What is the average daily activity pattern?

### 1. Makeing a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis) 


  first we start by finding average number of step per day for y-axis of plot
 
 
```{r, echo=TRUE}
avgstep<- activity %>% group_by(interval) %>% na.omit()  %>% summarise(totalstep=mean(steps))
head(avgstep)
```
 
 then we create a time series graph

```{r, echo=TRUE}



with(avgstep,plot(interval,totalstep,type = "l",  main = "The average daily activity pattern"))


```



### 2. Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?

The result is as follows 

``` {r, echo=TRUE}
avgstep[which.max(avgstep$totalstep),]

```
Interval 835 has the maximum number of steps which equals to 206

## Imputing missing values

### 1. Calculateion and reporting of the total number of missing values 

```{r, echo=TRUE}
sum(is.na(activity$steps))

```

 ### 2. Deviseing a strategy for filling in all of the missing values in the datase
 
My strategy here is to replace NA with the average of steps , to do so I will create a meansteps which contains the avergae value of steps.
 
```{r, echo=TRUE}

meansteps=mean(activity$step,na.rm=TRUE)

```
 
### 3. Create a new dataset that is equal to the original dataset but with the missing data filled in.

To prevent any future problem and to compare the diffrence between imputed and not impluted dataset I will create a copy of activity data set and will call it stepdata. Afterward, I will continue by replacing the missing values with the average step value, calculated in previous chunck.
 
 
```{r, echo=TRUE}

stepdata<-activity

stepdata$steps<-replace(stepdata$steps, is.na(stepdata$steps), meansteps)
        

```
 
Finally, I check if the new dataset have missing values 
 
```{r, echo=TRUE}
sum(is.na(stepdata$steps))

```
 
 Taaa Daaa everything is good 
 

 ### 4. Calculate Histogram,mean,median total number of step per day  for new data.frame
 
 a) Histogram of the new data 
```{r, echo=TRUE}

newtotalstep<-tapply(stepdata$steps,stepdata$date,sum)

newtotalstep<- stepdata %>% group_by(date) %>% summarise(totalstep=sum(steps))


hist(newtotalstep$totalstep, xlab = "number of steps",
      main = "total number of steps taken per day", col="green")

```

b) new mean and median are as follow 


```{r, echo=TRUE}

mean(newtotalstep$totalstep,na.rm=TRUE)
median(newtotalstep$totalstep,na.rm=TRUE)

```

c) diffrence between the new and old mean and median 

```{r, echo=TRUE}

meandif<- mean(totalstep,na.rm=TRUE)-mean(newtotalstep$totalstep,na.rm=TRUE)
meddif<-median(totalstep,na.rm=TRUE)-median(newtotalstep$totalstep,na.rm=TRUE)

paste("diffrence in mean is", meandif)

paste("diffrence in median is", meddif)

```

As  we can see the mean is the same as when we didn't substitue the NA variable,but the median has slightly changed. our new median is equall to are mean.

## Are there differences in activity patterns between weekdays and weekends?


### 1. Create a new factor variable in the dataset with two levels -- "weekday" and "weekend" indicating whether a given date is a weekday or weekend day.

Usin g the weekday function I will frist create a new charachter variable weekday which containt the weekday name.
Then using this weekday variable I will create a factor variable called day indicator which indicate wether that day is weekend or weekday

```{r, echo=TRUE}

stepdata$weekday<-weekdays(stepdata$date)
stepdata$dayindicator<-ifelse(stepdata$weekday %in% c("Sunday","Saturday"),0,1)
stepdata$dayindicator<- ordered(stepdata$dayindicator,
                          levels = c(0,1),
                          labels = c("Weekend","Weekday"))

head(stepdata,3)

```


### 2. Make a panel plot containing a time series plot 

 a. build the  new variable avrage step per day
 
```{r, echo=TRUE}
  
panelstepdata<- stepdata %>% group_by(interval,dayindicator) %>% summarise(tSteps=mean(steps))

head(panelstepdata,3)
  
```

  two time series graph for weekday no weekend 


```{r, echo=TRUE}

g<- ggplot(panelstepdata, aes(interval,tSteps,panelstepdata))
g+geom_line(col= "darkblue")+facet_grid(dayindicator~.)+labs(y= "Number of steps")+
theme(panel.grid.major = element_blank(),panel.grid.minor = element_blank())

```




